Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-05-05 paper.qmd
merge

$$
\varepsilon_k = \frac{s_{a} a_{h} + s_{d_k} a_{v}}{s_{d_k}}
$$ {#eq-effective-exposed-area}

@@

* constant debris fall speed $s_{d_k} = \qty{145}{mph} = S$ 
  ([we adopt this value from pag 7-8 in @patera2008])
* constant aircraft speed, $s_{a} = C_t$, equal to its cruise speed, under the
  simplification that the majority of the flight time spent cruising, 
* aircraft area as viewed from the top, $a_T = W_t \times L_t = A_t$, 
  where $W_t$ is the wingspan, and $L_t$ the length for the specific aircraft
  type $t$
* aircraft area as viewed from the front, $a_F = W_t \times H_t = F_t$, 
  where $W_t$ is the wingspan, and $H_t$ the height for the specific aircraft
  type $t$

@@

$$
\varepsilon_t = \frac{C_{t} F_{t} + S A_{t}}{S}
$$ {#eq-effective-exposed-area-simplified}

These aircraft properties are well-known design parameters. 
While there does not exist a single openly available date base for these aircraft parameters, we collected the information from a series of sources and integrated them into a lookup dataset.
The values $W_t$, $H_t$, $L_t$ and $C_{t}$ have been retrieved from [@performance_database;@skybrary;@doc8643]; we used $\varepsilon_t = \qty{1000}{m^2}$ missing entries or unknown aircraft types.

@@

* $E_{h,H}$ is the collision expectation in the hexagon $h$ at hour $H$,
* $W(h)$ is the median value of the weighting function over the area of hexagon
$h$,
* $\rho_{t,H}(h)$ is the traffic density for the \acr{ICAO} aircraft type $t$ in
cell $h$ at hour $H$, and
* $\varepsilon_t$ is effective exposed area for aircraft of type $t$.

@@

In few days (at the time of writing of this paper) at 23:06 May 10, 2025
($\pm 1.8$ days), the KOSMOS 482 Descent Craft should re-enter the
Earth's atmosphere [@kosmos482reentry].
Supposedly the lander module from a 1972 failed USSR Venera mission [@SpaceReviewKosmos],
it is reported to weigh about $\qty{480}{kg}$ and to have a diameter of approximately $\qty{1}{m}$.
Being built to survive Venus's atmosphere which is much denser and hotter than Earth's one (surface pressure 93 times that of Earth and temperatures reaching \ang{467}C (\ang{872}F)),
it could re-enter Earth intact and not break into smaller fragments some of which could burn up.

@@

Definition of Keplerian elements. The orbital elements describing the
trajectory of orbiting object are illustrated in association with an
inertial reference frame $\hat{I}$-$\hat{J}$-$\hat{K}$. The $I$-axis,
$\hat{I}$, points toward the vernal equinox and the $J$-axis, $\hat{J}$,
is perpendicular to the $I$-axis in the equatorial plane n the sense of
Earth's rotation with the $K$-axis pointing up. The semi-major axis,
$a$, the eccentricity, $e$, the inclination, $i$, the right ascension of
the ascending node, $\Omega$ , and the argument of perigee, $\omega$,
constitute, together with the mean anomaly at the initial position,
$M_0$, a set of six constants of motion that are preserved in the
absence of perturbation. True anomaly, $\nu$, specifies the location of
an orbiting object within the orbit. It is the angle between perigee and
the orbiting object's position vector measured in the direction of the
orbiting object's motion. Figures from [@sellers2003understanding],
respectively Figure 5-2 and Figure 5-9.
:::

@@

The angular position, $\theta$, of the re-entry point from the ascending
node in the orbit plane is $\theta = \omega + \nu$. It is related to the
latitude, $\varphi$, via the Cartesian coordinate $k$,
$k = R \sin(\varphi) = R \sin(\theta)\sin(i)$, or

@@

>>>>

$$
\varepsilon_k = \frac{s_{a} a_{F} + s_{d_k} a_{T}}{s_{d_k}}
$$ {#eq-effective-exposed-area}

@@

* constant debris fall speed $s_{d_k} = \qty{145}{mph} = S$ 
  ([we adopt this value from pag 7-8 in @patera2008])
* constant aircraft speed, $s_{a} = C_t$, equal to its cruise speed, under the
  simplification that the majority of the flight time spent cruising, 
* aircraft area as viewed from the top, $a_T = W_t \times L_t = T_t$, 
  where $W_t$ is the wingspan, and $L_t$ the length for the specific aircraft
  type $t$
* aircraft area as viewed from the front, $a_F = W_t \times H_t = F_t$, 
  where $W_t$ is the wingspan, and $H_t$ the height for the specific aircraft
  type $t$

@@

$$
\varepsilon_t = \frac{C_{t} F_{t} + S T_{t}}{S}
$$ {#eq-effective-exposed-area-simplified}

These aircraft properties are well-known design parameters. 
While there does not exist a single openly available date base for these aircraft parameters, we collected the information from a series of sources and integrated them into a lookup dataset.
The values $W_t$, $H_t$, $L_t$ and $C_{t}$ have been retrieved from [@performance_database;@skybrary;@doc8643]; we used $\varepsilon_t = \qty{500}{m^2}$
and $pax = 7$ for missing entries which usually are due to smaller aircraft not listed in the above sources.

@@

* $E_{h,H}$ is the collision expectation in the hexagon $h$ at hour $H$,
* $W(h)$ is the mean value of the weighting function over the area of hexagon
$h$,
* $\rho_{t,H}(h)$ is the traffic density for the \acr{ICAO} aircraft type $t$ in
cell $h$ at hour $H$, and
* $\varepsilon_t$ is effective exposed area for aircraft of type $t$.

@@

In few days (at the time of writing of this paper) at 23:06 May 10, 2025
($\pm 1.8$ days), the KOSMOS 482 Descent Craft should re-enter the
Earth's atmosphere [@kosmos482reentry].
This is the lander module from a 1972 failed USSR Venera mission [@SpaceReviewKosmos],
it is reported to weigh about $\qty{480}{kg}$ and to have a diameter of approximately $\qty{1}{m}$.
Being built to survive Venus's atmosphere which is much denser and hotter than Earth's one (surface pressure 93 times that of Earth and temperatures reaching \ang{467}C (\ang{872}F)),
it could re-enter Earth intact and not break into smaller fragments some of which could burn up.

@@

Definition of Keplerian elements. The orbital elements describing the
trajectory of orbiting object are illustrated in association with an
inertial reference frame $\hat{I}$-$\hat{J}$-$\hat{K}$. The $I$-axis,
$\hat{I}$, points toward the vernal equinox and the $J$-axis, $\hat{J}$,
is perpendicular to the $I$-axis in the equatorial plane n the sense of
Earth's rotation with the $K$-axis pointing up. The semi-major axis,
$a$, the eccentricity, $e$, the inclination, $i$, the right ascension of
the ascending node, $\Omega$, and the argument of perigee, $\omega$,
constitute, together with the mean anomaly at the initial position,
$M_0$, a set of six constants of motion that are preserved in the
absence of perturbation. True anomaly, $\nu$, specifies the location of
an orbiting object within the orbit. It is the angle between perigee and
the orbiting object's position vector measured in the direction of the
orbiting object's motion. Figures from [@sellers2003understanding],
respectively Figure 5-2 and Figure 5-9.
:::

@@

The angular position, $\theta$, of the re-entry point from the ascending
node[^ascending-node] in the orbit plane is $\theta = \omega + \nu$. It is related to the
latitude, $\varphi$, via the Cartesian coordinate $k$,
$k = R \sin(\varphi) = R \sin(\theta)\sin(i)$, or

@@

[^ascending-node]: For an earth satellite, the ascending node is the point in its orbit where a satellite crosses the equatorial plane going from south to north.
